---
title: "Credit Risk V1"
output: html_document
date: "2023-11-13"
---
# import all packages
```{r}
library(dplyr)
library(dummy)
library(FNN)
library(class)
library(caret)
library(ggplot2)
library(lattice)
library(kknn)
library(ROCR)
```

# convert cb_person_default_on_file col into factor
```{r}
options(repos = c(CRAN = "https://cloud.r-project.org/"))
set.seed(1)
creditData=na.omit(read.csv ("credit_risk_dataset.csv", header = TRUE))
num_samples = dim(creditData)[1]

creditData$cb_person_default_on_file <- ifelse(creditData$cb_person_default_on_file == "Y", 1, 0)
creditData$cb_person_default_on_file <- factor(creditData$cb_person_default_on_file, levels = c(1, 0))

head(creditData)
```

# getting an overview of classification variables
```{r}
home_ownership_types <- creditData %>% 
  count(person_home_ownership)

print(home_ownership_types)

loan_intent <- creditData %>% 
  count(loan_intent)

print(loan_intent)

loan_grade <- creditData %>% 
  count(loan_grade)

print(loan_grade)

cb_person_default_on_file <- creditData %>% 
  count(cb_person_default_on_file)

print(cb_person_default_on_file)

```
```{r}
## loan status is Y 
summary (creditData$person_home_ownership)



# creating dummy variables
temp_ownership <- data.frame(person_home_ownership = creditData$person_home_ownership)
creditData <- cbind(creditData, dummy(temp_ownership))
creditData <- subset(creditData, select = -person_home_ownership)

temp_intent <- data.frame(intent = creditData$loan_intent)
creditData <- cbind(creditData, dummy(temp_intent))
creditData <- subset(creditData, select = -loan_intent)

temp_grade <- data.frame(grade = creditData$loan_grade)
creditData <- cbind(creditData, dummy(temp_grade))
creditData <- subset(creditData, select = -loan_grade)

creditData <- as.data.frame(sapply(creditData, function(x) as.numeric(as.character(x))))
creditData <- creditData


# normalizing the data
creditData$person_age = (creditData$person_age-mean(creditData$person_age))/sd(creditData$person_age)
creditData$person_income = (creditData$person_income - mean(creditData$person_income))/sd(creditData$person_income)
creditData$person_emp_length = (creditData$person_emp_length - mean(creditData$person_emp_length))/sd(creditData$person_emp_length)
creditData$loan_amnt = (creditData$loan_amnt - mean(creditData$loan_amnt))/sd(creditData$loan_amnt)
creditData$loan_int_rate = (creditData$loan_int_rate - mean(creditData$loan_int_rate))/sd(creditData$loan_int_rate)
creditData$cb_person_cred_hist_length = (creditData$cb_person_cred_hist_length - mean(creditData$cb_person_cred_hist_length))/sd(creditData$cb_person_cred_hist_length)
```

# K fold cross validation to deterine the best k
```{r}
bestK <- 0
minerrors <- Inf

# Initialize the progress bar
pbcount <- 0
pb <- txtProgressBar(min = 0, max = 100, style = 3)

for(iter in 55:65)
{
  AllErrors=numeric(50)
  for(fold in 1:10)
  {
    #Get Training at Testing sets
    training <- sample(1:num_samples, 0.8 * num_samples, replace=FALSE)
    trainingSet <- subset(creditData[training, ]) 
    testing <- setdiff(1:num_samples,training) 
    testingSet <- subset(creditData[testing, ])
    # Get the features of the training set
    trainingfeatures <- subset(trainingSet, select=c(-loan_status))
    # Get the labels of the training set
    traininglabels <- trainingSet$loan_status
    # Get the features of the testing set
    testingfeatures <- subset(testingSet, select=c(-loan_status))
    
    sizeTestSet = dim(testingSet)[1]
    knnModel = kknn(loan_status ~. ,train = trainingSet,test = testingSet,k=iter)
    predProbs <- fitted(knnModel)
    predictions = ifelse(predProbs > 0.5, 1, 0)
    
    error = sum(predictions != testingSet$loan_status)
    mis_rate=error/sizeTestSet
    AllErrors[fold] = mis_rate
    
    # Setting the progress bar
    pbcount <- pbcount + 1
    setTxtProgressBar(pb, pbcount)
  }
  
  currError = mean(AllErrors)
  if (currError < minerrors){
    bestK = iter
    minerrors = currError
  }
}
close(pb)
print(bestK)
```
# Splitting training and testing set
```{r}
num_samples = dim(creditData)[1]
training=sample(1:num_samples,0.8*num_samples,replace=FALSE) 
trainingsSet=subset(creditData[training,])

testing = setdiff(1:num_samples, training)
testingSet = subset(creditData[testing, ])
```
# using loan status as training and testing features
```{r}
trainingfeature = subset(trainingsSet, select=c(-loan_status))

traininglabel =trainingsSet$loan_status

testingfeature = subset(testingSet, select=c(-loan_status))
```
# train the model with the traiing set
```{r}
knnModel <- kknn(loan_status ~. , train = trainingSet, test = testingSet, k=bestK)

predProbs <- fitted(knnModel)
predictionLabels = ifelse(predProbs > 0.5, 1, 0)

head(predictionLabels)
print(class(predictionLabels))
head(testingSet$loan_status)
```
# make predictions with the testing set
```{r}
sizeTestSet = dim(testingSet)[1] 
error = sum(predictionLabels!=testingSet$loan_status)

mis_rate=error/sizeTestSet

print(mis_rate)
```
# generate the confusion matrix
```{r}
conf_matrix <- confusionMatrix(as.factor(predictionLabels), as.factor(testingSet$loan_status), positive = "1")
print(conf_matrix)
```
# plot the roc and auc to determine the performance of the model
```{r}
pred = prediction(predProbs, testingSet$loan_status)
pred_knn = pred
print(predProbs)
roc_obj = performance(pred, measure = "tpr", x.measure = "fpr")
auc_obj = performance(pred, measure = "auc")
plot(roc_obj, main = paste("Area under the curve:", round(auc_obj@y.values[[1]], 2)))
```