---
title: "Loan Logistic Regression"
output: html_document
date: "2023-11-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load the relevant packages
```{r}
library(dplyr)
library(caret)
library(ROCR)
library(car)
library(dummy)
library(fastDummies)
```

# Read the data into R and remove all "NA" rows
```{r}
LoanData = read.csv("credit_risk_dataset.csv")
head(LoanData)
LoanData = na.omit(LoanData)
```

# Remove loan_percent_income becuase it's a redundant variable - the dataset already has loan amount and income 
```{r}
LoanData$loan_percent_income = NULL
```

# Convert categorical data into dummy variables and the default on file into a binary variable 
```{r}
LoanData = dummy_cols(LoanData, select_columns = "person_home_ownership")
LoanData = dummy_cols(LoanData, select_columns = "loan_intent")
LoanData = dummy_cols(LoanData, select_columns = "loan_grade")
default_on_file = ifelse(LoanData$cb_person_default_on_file == "Y", 1, 0)
LoanData$person_home_ownership = NULL
LoanData$loan_intent = NULL
LoanData$loan_grade = NULL
LoanData$cb_person_default_on_file = NULL
head(LoanData)
```

# Set the seed so results can be replicated 
```{r}
set.seed(4469)
```

# Split the dataset into training set (80%) and test test (20%)
```{r}
splitData = caret::createDataPartition(LoanData$loan_status, p = 0.8, list=FALSE)
trainData = LoanData[splitData,]
testData = LoanData[-splitData,]
```

# Apply Logistic Regression on Training set
```{r}
trainDataLR = glm(loan_status ~ ., data=trainData, family=binomial(link="logit"))
summary(trainDataLR)
```


# Retrain the data after removing th einsignificnat variables 
```{r}
trainDataLR = glm(loan_status ~ person_age  + person_income + person_emp_length + loan_amnt + loan_int_rate + person_home_ownership_MORTGAGE + person_home_ownership_OWN + loan_intent_DEBTCONSOLIDATION + loan_intent_EDUCATION + loan_intent_HOMEIMPROVEMENT + loan_intent_MEDICAL + loan_intent_PERSONAL, data=trainData, family=binomial(link="logit"))
summary(trainDataLR)
```

# Predicting probabilities and set threshold at 0.5
```{r}
pred_probs = predict(trainDataLR, testData, type = "response")
testData$Predicted = ifelse(pred_probs > 0.5, 1, 0)
```

# Convert Predicted and Churned columns to factor with explicit levels
```{r}
testData$Predicted = factor(testData$Predicted, levels = c(0, 1))
testData$loan_status = factor(testData$loan_status, levels = c(0, 1))
head(testData)
```

# Confusionmatrix; its values depend on the treshold we choose (0.5 in this case)
```{r}
confusionMatrix(testData$Predicted, testData$loan_status, positive = "1")
```

# Calculating and plotting ROC curve
```{r}
pred = prediction(pred_probs, testData$loan_status)
pred_logistic = pred
roc_obj = performance(pred, measure = "tpr", x.measure = "fpr")
auc_obj = performance(pred, measure = "auc")
plot(roc_obj, main = paste("Area under the curve:", round(auc_obj@y.values[[1]], 2)))
```

